---
title: "100-references"
output: html_document
date: "2024-02-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
```

We are looking at analyzing RADseq data from Siphateles spp.    

There are several possible reference genomes from Siphateles under construction with the CCGP. Existing public genomes with divergence time from Siphateles:     


_Chrosomus eos_ 33 mya.  GCA_036326025.1 is chromosomal-level.    
_Gila orcutti_ 12.1 mya GCA_026230005.1 506 contigs, 179 scaffolds with N50 of 50.5 Mb, 1.3 Gb total.   
_Gila robusta_ 12.1 mya GCA_030770095.1 2,618 contigs,  level assembly.    
_Pimephales promelas_ 43 mya GCF_016745375.1 1.1 Gb scaffolded assembly

Some options for reference genomes, here is a study using a variety of references https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6356778/.    


_Gila orcutti_ 12.1 mya GCA_026230005.1 506 contigs, 179 scaffolds with N50 of 50.5 Mb, 1.3 Gb total.   
