---
title: "108-recalled-snps"
output: html_document
date: "2024-05-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(phangorn)
library(ggtree)
library(ape)
library(tanggle)
```


Had good results with the 532 data set for treemix. Will recall, after some filtering for sites in obesus, including Pogonichthys and Cowhead Lake fish.

Note to self to double check pogonichthys.

So, 543 inds +  G. orcuttii P. macrolepidotus, then filtering for small sample sizes.

```{r}
calls<-read_csv("meta/phylo-582.csv") %>% 
  filter(`Tributary/Collection Location` != "3/8 Mile Spring")
```

582 inds w/o "3/8 Mile Spring".   

```{r}
clst<-calls %>% mutate(Inter=GVL_Code) %>% mutate(Cluster=paste0(Species,"_",`Tributary/Collection Location`)) 
clst$Cluster<-gsub(" ","_",clst$Cluster)

clst<-clst %>% filter(!Cluster %in% c("obesus_Little_Hot_Creek_Pond","obesus_Little_Hot_Creek","obesus_Upper_Owens_Gorge","obesus_SE_Pond_(White_Mountain_Research_Center",
                                      "obesus_SW_Pond_(White_Mountain_Research_Center")) %>% group_by(Cluster) %>% mutate(Count=n())
clst1<-filter(clst, Count > 2)
clst<-clst %>% filter(Cluster=="G._orcuttii_Desert_Discovery_Pond") %>% bind_rows(clst1)
clst %>% summarize(Count=n())

clst2<-clst %>% mutate(Intra="1")
clst2$Sample<-gsub("_1","",clst2$Sample)
clst2
```

```{r}
clst2 %>% write_csv("meta/544.csv")
clst2 %>% dplyr::ungroup() %>% dplyr::select(Sample) %>% write_tsv(file="meta/544.names", col_names=FALSE)
clst2 %>% dplyr::ungroup() %>% dplyr::select(Path2) %>% write_tsv(file="bamlists/544.bamlist", col_names=FALSE)
```

95% thresh, 544*.95
```{sh, eval=FALSE}
srun -t 36:00:00 -p bigmemh --mem=128G --nodes=1 --ntasks-per-node=1 --cpus-per-task=12 $HOME/angsd/angsd -nthreads 24 \
-minInd 517 -bam bamlists/544.bamlist -ref /home/maccamp/genomes/gila-orcutti/GCA_026230005.1_fGilOrc1.0.hap1_genomic.fna.gz \
-out outputs/108/snps-544-01  \
-minMaf 0.01 -minMapQ 10 -minQ 20 -GL 1 -doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doGeno 2 -doPost 1 -postCutoff 0.95 -doPlink 2  > outputs/108/snps-544-01.out 2> outputs/108/snps-544-01.err &
  
#90 thresh
srun -t 36:00:00 -p bigmemh --mem=128G --nodes=1 --ntasks-per-node=1 --cpus-per-task=12 $HOME/angsd/angsd -nthreads 24 \
-minInd 490 -bam bamlists/544.bamlist -ref /home/maccamp/genomes/gila-orcutti/GCA_026230005.1_fGilOrc1.0.hap1_genomic.fna.gz \
-out outputs/108/snps-544-01-90  \
-minMaf 0.01 -minMapQ 10 -minQ 20 -GL 1 -doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doGeno 2 -doPost 1 -postCutoff 0.95 -doPlink 2  > outputs/108/snps-544-01-90.out 2> outputs/108/snps-544-01-90.err &

#Altering map q, but 95% thresh
srun -t 36:00:00 -p bigmemh --mem=128G --nodes=1 --ntasks-per-node=1 --cpus-per-task=12 $HOME/angsd/angsd -nthreads 24 \
-minInd 517 -bam bamlists/544.bamlist -ref /home/maccamp/genomes/gila-orcutti/GCA_026230005.1_fGilOrc1.0.hap1_genomic.fna.gz \
-out outputs/108/snps-544-01-q  \
-minMaf 0.01 -minMapQ 20 -minQ 20 -GL 1 -doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doGeno 2 -doPost 1 -postCutoff 0.95 -doPlink 2  > outputs/108/snps-544-01q.out 2> outputs/108/snps-544-01q.err &


```

	[ALL done] walltime used =  28519.00 sec


Should be able root at Pogonichthys and run a SVDquartets analysis.

Subsequently, we can filter and run a treemix analysis (filter means removing Gila orcuttii and Pognonichthys and filtering with BCFTools)

Convert to VCF and prune

250506 variants
306323 variants 90%.  
247384 variants with q20 & 95%
It seems like the initial snp Calling has a lot to do with how things turn out, no matter how the MAF is manipulated. 


```{sh, eval=FALSE}
plink --tped snps-544-01.tped --tfam snps-544-01.tfam  --out plink-binary --recode --allow-extra-chr --noweb
plink --ped plink-binary.ped --map plink-binary.map --recode vcf --allow-extra-chr -out plink
bgzip plink.vcf 
tabix plink.vcf.gz

plink --tped snps-544-01-90.tped --tfam snps-544-01-90.tfam  --out plink-binary-90 --recode --allow-extra-chr --noweb
plink --ped plink-binary-90.ped --map plink-binary-90.map --recode vcf --allow-extra-chr -out plink-90
bgzip plink-90.vcf 
tabix plink-90.vcf.gz

plink --tped snps-544-01-q.tped --tfam snps-544-01-q.tfam  --out plink-binary-q --recode --allow-extra-chr --noweb
plink --ped plink-binary-q.ped --map plink-binary-q.map --recode vcf --allow-extra-chr -out plink-q
bgzip plink-q.vcf 
tabix plink-q.vcf.gz

#renaming vcf with sample names, filtering again on MAF, and then pruning, selecting random SNP

module load bcftools/1.13

#Basically, pruning before was like this: bcftools +prune -l 0.20 -w 10000
bcftools reheader --samples meta/544.names outputs/108/plink.vcf.gz | bcftools +fill-tags -- -t MAF | bcftools view -q 0.01:minor | bcftools +prune -m 0.2 > outputs/108/pruned-02.vcf
#73002 snps



bcftools reheader --samples meta/544.names outputs/108/plink.vcf.gz | bcftools +fill-tags -- -t MAF | bcftools view -q 0.01:minor | bcftools +prune -m 0.2 -w 10000bp -n 1 -N rand > outputs/108/pruned.vcf
# 11038 snps

#Making a high MAF
bcftools reheader --samples meta/544.names outputs/108/plink.vcf.gz | bcftools +fill-tags -- -t MAF | bcftools view -q 0.3:minor | bcftools +prune -m 0.2 -w 10000bp -n 1 -N rand > outputs/108/pruned-maf03.vcf
#3207 SNPS, but still seems to have a good structure, reticulations present between mohavensis nad the others. So, I think we can test something like this in treemix. 

bcftools reheader --samples meta/544.names outputs/108/plink-90.vcf.gz | bcftools +fill-tags -- -t MAF | bcftools view -q 0.3:minor | bcftools +prune -m 0.2 -w 10000bp -n 1 -N rand > outputs/108/pruned-90-maf03.vcf
# 4018 snps
bcftools reheader --samples meta/544.names outputs/108/plink-90.vcf.gz | bcftools +fill-tags -- -t MAF | bcftools view -q 0.01:minor | bcftools +prune -m 0.2 -w 10000bp -n 1 -N maxAF > outputs/108/pruned-90-maf01.vcf
# 11978 snps


#drop -n 1 -n rand? Expand window size?
bcftools reheader --samples meta/544.names outputs/108/plink.vcf.gz | bcftools +fill-tags -- -t MAF | bcftools view -q 0.01:minor | bcftools +prune -m 0.2 -w 10000bp -n 1 -N maxAF > outputs/108/pruned-m2.vcf
# 87259 after dropping -n 1 -n rand that's a lot more snps!!
# n 1 -N maxAF, 11038 snps again, but, hopefully more poly


bcftools reheader --samples meta/544.names outputs/108/plink.vcf.gz | bcftools +fill-tags -- -t MAF | bcftools view -q 0.01:minor | bcftools +prune -m 0.3 -w 1000bp -n 1 -N maxAF > outputs/108/pruned-m3.vcf
#[default 100000kb]
#gives 79,993 snps at default windo
# givs 12,501 snps at -w1000bp -n 1 -N maxAF

bcftools reheader --samples meta/544.names outputs/108/plink.vcf.gz | bcftools +fill-tags -- -t MAF | bcftools view -q 0.01:minor | bcftools +prune -m 0.2 -w 100000bp -n 1 -N maxAF > outputs/108/pruned-100K.vcf
#554 5761


bcftools reheader --samples meta/544.names outputs/108/plink.vcf.gz | bcftools +fill-tags -- -t MAF | bcftools view -q 0.10:minor | bcftools +prune -m 0.3 -w 10000bp -n 1 -N rand > outputs/108/pruned-maf10.vcf
#m2 554 6508
#m3 554 6854


#Coming to terms with the static results....
bcftools reheader --samples meta/544.names outputs/108/plink-q.vcf.gz | bcftools +fill-tags -- -t MAF | bcftools view -q 0.01:minor | bcftools +prune -m 0.2 -w 10000bp -n 1 -N maxAF > outputs/108/pruned-q-maf01.vcf


```

Locally
```{sh, eval=FALSE}
# Convert to phylip 
conda activate py2;
~/github/mccloud-rrt/vcf2phylip.py -i pruned.vcf 
~/github/mccloud-rrt/vcf2phylip.py -i pruned-maf10.vcf 
~/github/mccloud-rrt/vcf2phylip.py -i pruned-100K.vcf 

~/github/mccloud-rrt/vcf2phylip.py -i pruned-m2.vcf 
~/github/mccloud-rrt/vcf2phylip.py -i pruned-m3.vcf 

~/github/mccloud-rrt/vcf2phylip.py -i pruned-maf03.vcf 
~/github/mccloud-rrt/vcf2phylip.py -i pruned-90-maf03.vcf
~/github/mccloud-rrt/vcf2phylip.py -i pruned-90-maf01.vcf
~/github/mccloud-rrt/vcf2phylip.py -i pruned-q-maf01.vcf

conda deactivate;
```

Try SVDQuartets. Confirmed P. macrolepidotus ID.     
        dimensions ntax = 554 nchar = 11038;

```{r}
m544<-clst2 %>% mutate(SVDQ=paste0(Sample,"T"))
m544 %>% select(SVDQ) %>%  summarize(Inds=paste0(SVDQ, collapse="\t"))  %>% write_tsv("outputs/108/block.tsv")
```


Create tree meta

```{r}
speciesm<-m544 %>% select(Cluster, Species, Subspecies, Vicinity, `Tributary/Collection Location`)
speciesm

```


```{r}
s<-read.nexus("outputs/108/svdq.tre")
s<-consensus(s, p=0.5)
s<-root(s, c("P._macrolepidotus_Cowhead_Lake"))

ggtree(s, branch.length = 'none') %<+% speciesm +
  geom_tippoint(aes(shape=Species, fill=Subspecies, size=Count)) + 
  geom_nodelab() + 
  geom_tiplab(size=4, offset = 0.5) +
  xlim(0,20) +
  scale_fill_viridis_d(option="magma", na.value="grey50") +
  guides(fill = guide_legend(override.aes = list(pch = 22))) +
  scale_shape_manual(values=c(25,23,25,23,24,22,21,23,25))



ggsave("outputs/108/svdq-test.pdf", width=11, height = 8.5)

```

# ML analysis
```{sh,eval=FALSE}
iqtree2 -s pruned.min4.phy -T AUTO -st DNA -m MFP+ASC -bb 10000 -nm 2000 -bcor 0.9 --redo
iqtree2 -s pruned.min4.phy.varsites.phy -T AUTO -st DNA -m MFP+ASC -bb 10000 -nm 2000 -bcor 0.9 --redo

iqtree2 -s pruned-maf03.min4.phy -T AUTO -st DNA -m MFP+ASC -bb 10000 -nm 2000 -bcor 0.9 --redo
iqtree2 -s pruned-maf03.min4.phy.varsites.phy -T AUTO -st DNA -m MFP+ASC -bb 10000 -nm 2000 -bcor 0.9 --redo

iqtree2 -s pruned-02.min4.phy -T AUTO -st DNA -m MFP+ASC -bb 10000 -nm 2000 -bcor 0.9 --redo
iqtree2 -s pruned-02.min4.phy.varsites.phy -T AUTO -st DNA -m MFP+ASC -bb 10000 -nm 2000 -bcor 0.9 --redo
# Only 45836 sites considered variant by iqtree

iqtree2 -s pruned-q-maf01.min4.phy -T AUTO -st DNA -m GTR+ASC -bb 10000 -nm 2000 -bcor 0.9 --redo
iqtree2 -s pruned-q-maf01.min4.phy.varsites.phy -T AUTO -st DNA -m GTR+ASC -bb 10000 -nm 2000 -bcor 0.9 --redo

```



## What about a network?

```{r}
dat<-read.dna(file="outputs/108/pruned.min4.phy")
write.nexus.data(dat, file="outputs/108/pruned.min4.nexus")

dat<-read.dna(file="outputs/108/pruned-maf10.min4.phy")
write.nexus.data(dat, file="outputs/108/pruned-maf10.min4.nexus")

dat<-read.dna(file="outputs/108/pruned-100K.min4.phy")
write.nexus.data(dat, file="outputs/108/pruned-100K.min4.nexus")

dat<-read.dna(file="outputs/108/pruned-maf03.min4.phy")
write.nexus.data(dat, file="outputs/108/pruned-maf03.min4.nexus")

dat<-read.dna(file="outputs/108/iqtree/pruned-02.min4.phy.varsites.phy")
write.nexus.data(dat, file="outputs/108/iqtree/pruned-02.nexus")

dat<-read.dna(file="outputs/108/pruned-90-maf03.min4.phy")
write.nexus.data(dat, file="outputs/108/pruned-90-maf03.min4.nexus")

dat<-read.dna(file="outputs/108/pruned-90-maf01.min4.phy")
write.nexus.data(dat, file="outputs/108/pruned-90-maf01.min4.nexus")

dat<-read.dna(file="outputs/108/iqtree/pruned-q-maf01.min4.phy.varsites.phy")
write.nexus.data(dat, file="outputs/108/pruned-q-maf01.min4.phy.varsites.nexus")

```

Made a network in SplitsTree

```{r}
#net<-read.nexus.networx("outputs/108/pruned-90-maf03.network")
#net<-read.nexus.networx("outputs/108/pruned-90-maf01.network")

net<-read.nexus.networx("outputs/108/pruned-90-maf01.network")

```


```{r}
g<-ggsplitnet(net)  
g$data<-left_join(g$data, m544, by=c("label"="Sample")) 

n<-g + 
  geom_tippoint(aes(fill=Species), cex=5, pch=21, alpha=1.0) +
  scale_fill_viridis_d(option="H") +
  guides(fill = guide_legend(override.aes = list(pch = 22)))

n

```



# Downample and recall?
Obesus is highly represented, and the placement of mohavensis is a bit odd I think.

```{r}
m429<-clst2 %>% filter(!`Tributary/Collection Location` %in% c("Lower Twin Lakes", "Pyramid Lake", "Walker Lake", "Topaz Lake","Upper Homboldt River","Soldier Meadows", "Turley Pond","Casey Pond","Stillwater Wildlife Refuge"))
write_csv(m429, "meta/429.csv")
m429 %>% dplyr::ungroup() %>% dplyr::select(Sample) %>% write_tsv(file="meta/429.names", col_names=FALSE)
m429 %>% dplyr::ungroup() %>% dplyr::select(Path2) %>% write_tsv(file="bamlists/429.bamlist", col_names=FALSE)
```

```{sh, eval=FALSE}
#Altering map q, but 95% thresh
srun -t 36:00:00 -p bigmemh --mem=128G --nodes=1 --ntasks-per-node=1 --cpus-per-task=12 $HOME/angsd/angsd -nthreads 24 \
-minInd 408 -bam bamlists/429.bamlist -ref /home/maccamp/genomes/gila-orcutti/GCA_026230005.1_fGilOrc1.0.hap1_genomic.fna.gz \
-out outputs/108/snps-429-01 \
-minMaf 0.01 -minMapQ 20 -minQ 20 -GL 1 -doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doGeno 2 -doPost 1 -postCutoff 0.95 -doPlink 2  > outputs/108/snps-429-01.out 2> outputs/108/snps-429-01.err &

#227006 variants

plink --tped snps-429-01.tped --tfam snps-429-01.tfam  --out plink-binary-429 --recode --allow-extra-chr --noweb
plink --ped plink-binary-429.ped --map plink-binary-429.map --recode vcf --allow-extra-chr -out plink-429
bgzip plink-429.vcf 
tabix plink-429.vcf.gz

#Using random seed: 1716420232
bcftools reheader --samples meta/429.names outputs/108/plink-429.vcf.gz | bcftools +fill-tags -- -t MAF | bcftools view -q 0.01:minor | bcftools +prune -m 0.2 -w 10000bp -n 1 -N rand > outputs/108/pruned-429-maf01.vcf

#local
conda activate py2; ~/github/mccloud-rrt/vcf2phylip.py -i pruned-429-maf01.vcf; conda deactivate; 

iqtree2 -s pruned-429-maf01.min4.phy -T AUTO -st DNA -m GTR+ASC -bb 10000 -nm 2000 -bcor 0.9 --redo
iqtree2 -s pruned-429-maf01.min4.phy.varsites.phy -T AUTO -st DNA -m GTR+ASC -bb 10000 -nm 2000 -bcor 0.9 --redo
# 7185 sites
```



```{r}
dat<-read.dna(file="outputs/108/pruned-429-maf01.min4.phy.varsites.phy")
write.nexus.data(dat, file="outputs/108/pruned-429.nexus")
```

```{r}
net<-read.nexus.networx("outputs/108/pruned-429.network")
```


```{r}
g<-ggsplitnet(net)  
g$data<-left_join(g$data, m429, by=c("label"="Sample")) 

n<-g + 
  geom_tippoint(aes(fill=Species), cex=5, pch=21, alpha=1.0) +
  scale_fill_viridis_d(option="H") +
  guides(fill = guide_legend(override.aes = list(pch = 22)))

n

```

Can also downsample 'snyderi' lineage...  

#NJ

```{r}
nj<-read.nexus("outputs/108/pruned-429.tre")
nj<-midpoint(nj)
ggtree(nj) %<+% relocate(m429,Sample) +
    geom_tippoint(aes(fill=Species), pch=21) +
  geom_tiplab(align = TRUE, size=1, aes(label=`Tributary/Collection Location`))
ggsave("outputs/108/nj.tre.pdf", width=14, height=20)
```

# Downsample snyderi

```{r}
m386<-clst2 %>% filter(!`Tributary/Collection Location` %in% c("Lower Twin Lakes", "Pyramid Lake", "Walker Lake", "Topaz Lake","Upper Homboldt River","Soldier Meadows", "Turley Pond","Casey Pond","Stillwater Wildlife Refuge","NE Pond (White Mountain Research Center)","Cottonwood Pond (White MT Research Center)"))


write_csv(m386, "meta/386.csv")
m386 %>% dplyr::ungroup() %>% dplyr::select(Sample) %>% write_tsv(file="meta/386.names", col_names=FALSE)
m386 %>% dplyr::ungroup() %>% dplyr::select(Path2) %>% write_tsv(file="bamlists/386.bamlist", col_names=FALSE)
```


```{sh, eval=FALSE}
#Altering map q, but 95% thresh
srun -t 36:00:00 -p bigmemh --mem=128G --nodes=1 --ntasks-per-node=1 --cpus-per-task=12 $HOME/angsd/angsd -nthreads 24 \
-minInd 408 -bam bamlists/386.bamlist -ref /home/maccamp/genomes/gila-orcutti/GCA_026230005.1_fGilOrc1.0.hap1_genomic.fna.gz \
-out outputs/108/snps-386-01 \
-minMaf 0.01 -minMapQ 20 -minQ 20 -GL 1 -doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doGeno 2 -doPost 1 -postCutoff 0.95 -doPlink 2  > outputs/108/snps-386-01.out 2> outputs/108/snps-386-01.err &

#

plink --tped snps-386-01.tped --tfam snps-386-01.tfam  --out plink-binary-386 --recode --allow-extra-chr --noweb
plink --ped plink-binary-386.ped --map plink-binary-386.map --recode vcf --allow-extra-chr -out plink-386
bgzip plink-386.vcf 
tabix plink-386.vcf.gz

#Using random seed: 
bcftools reheader --samples meta/386.names outputs/108/plink-386.vcf.gz | bcftools +fill-tags -- -t MAF | bcftools view -q 0.01:minor | bcftools +prune -m 0.2 -w 10000bp -n 1 -N rand > outputs/108/pruned-386-maf01.vcf

#local
conda activate py2; ~/github/mccloud-rrt/vcf2phylip.py -i pruned-386-maf01.vcf; conda deactivate; 

iqtree2 -s pruned-386-maf01.min4.phy -T AUTO -st DNA -m GTR+ASC -bb 10000 -nm 2000 -bcor 0.9 --redo
iqtree2 -s pruned-386-maf01.min4.phy.varsites.phy -T AUTO -st DNA -m GTR+ASC -bb 10000 -nm 2000 -bcor 0.9 --redo
#  sites
```