---
title: "300-newarkensis-isolatus"
output: html_document
date: "2026-01-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(pcadapt)
library(ape)
library(grid)
library(ggpubr)
library(ggrepel)
library(ggtree)
```

Is there any phylogenetic structure to S. newarkensis?

```{r}
meta<-read_csv("meta/newarkensis-isolatus-500K-meta.csv")
meta %>% select(Path) %>% write_tsv("bamlists/341.bamlist", col_names = FALSE)
meta %>% select(GVL_Code) %>% write_tsv("bamlists/341.names", col_names = FALSE)

nmeta<-meta %>% filter(Species !="isolatus")
nmeta %>% select(Path) %>% write_tsv("bamlists/321.bamlist", col_names = FALSE)
nmeta %>% select(GVL_Code) %>% write_tsv("bamlists/321.names", col_names = FALSE)
```

Call snps at 95% threshold with a minMaf of 0.01.   
	-> Number of sites retained after filtering: 76144 


sinfo needed
```{sh, eval=FALSE}
srun -t 36:00:00 -p bmh --mem=128G --nodes=1 --ntasks-per-node=1 --cpus-per-task=12 $HOME/angsd/angsd -nthreads 24 \
-minInd 324 -bam bamlists/341.bamlist -ref /home/maccamp/genomes/gila-orcutti/GCA_026230005.1_fGilOrc1.0.hap1_genomic.fna.gz \
-out outputs/300/snps-341-01  \
-minMaf 0.01 -minMapQ 20 -minQ 20 -GL 1 -doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doGeno 2 -doPost 1 -postCutoff 0.95 -doPlink 2  > outputs/300/snps-341-01.out 2> outputs/300/snps-341-01.err &

#convert to .vcf
plink --tped snps-341-01.tped --tfam snps-341-01.tfam  --out plink-binary --recode --allow-extra-chr --noweb
plink --ped plink-binary.ped --map plink-binary.map --recode vcf --allow-extra-chr -out plink
bgzip plink.vcf 
tabix plink.vcf.gz

bcftools reheader --samples bamlists/341.names outputs/300/plink.vcf.gz  > outputs/300/renamed.vcf.gz
tabix outputs/300/renamed.vcf.gz

bcftools view -S bamlists/321.names outputs/300/renamed.vcf.gz > outputs/300/newarkensis.vcf

#prune on farm
module load bcftools
bcftools reheader --samples bamlists/341.names outputs/300/plink.vcf.gz | bcftools +fill-tags -- -t MAF | bcftools view -q 0.01:minor |  bcftools +prune -m 0.2 -w 10000bp -n 1 -N maxAF > outputs/300/pruned01-max.vcf

bcftools reheader --samples bamlists/341.names outputs/300/plink.vcf.gz | bcftools +fill-tags -- -t MAF | bcftools view -q 0.01:minor |  bcftools +prune -m 0.2 -w 10000bp -n 1 -N rand > outputs/300/pruned01-rand.vcf

bcftools reheader --samples bamlists/341.names outputs/300/plink.vcf.gz | bcftools +fill-tags -- -t MAF | bcftools view -q 0.05:minor |  bcftools +prune -m 0.2 -w 10000bp -n 1 -N maxAF > outputs/300/pruned05-max.vcf

bcftools reheader --samples bamlists/341.names outputs/300/plink.vcf.gz | bcftools +fill-tags -- -t MAF | bcftools view -q 0.05:minor |  bcftools +prune -m 0.2 -w 10000bp -n 1 -N rand > outputs/300/pruned05-rand.vcf

bcftools reheader --samples bamlists/341.names outputs/300/plink.vcf.gz | bcftools +fill-tags -- -t MAF | bcftools view -q 0.10:minor |  bcftools +prune -m 0.2 -w 10000bp -n 1 -N maxAF > outputs/300/pruned10-max.vcf

bcftools reheader --samples bamlists/341.names outputs/300/plink.vcf.gz | bcftools +fill-tags -- -t MAF | bcftools view -q 0.10:minor |  bcftools +prune -m 0.2 -w 10000bp -n 1 -N rand > outputs/300/pruned10-rand.vcf

# get newarkensis
bcftools +fill-tags outputs/300/newarkensis.vcf -- -t MAF |  bcftools view -q 0.05:minor |  bcftools +prune -m 0.2 -w 10000bp -n 1 -N rand > outputs/300/newarkensis-pruned05-rand.vcf 

bcftools +fill-tags outputs/300/newarkensis.vcf -- -t MAF |  bcftools view -q 0.05:minor |  bcftools +prune -m 0.2 -w 10000bp -n 1 -N maxAF > outputs/300/newarkensis-pruned05-max.vcf

bcftools +fill-tags outputs/300/newarkensis.vcf -- -t MAF |  bcftools view -q 0.10:minor |  bcftools +prune -m 0.2 -w 10000bp -n 1 -N rand > outputs/300/newarkensis-pruned10-rand.vcf 

bcftools +fill-tags outputs/300/newarkensis.vcf -- -t MAF |  bcftools view -q 0.10:minor |  bcftools +prune -m 0.2 -w 10000bp -n 1 -N maxAF > outputs/300/newarkensis-pruned10-max.vcf

#trial phylogeny, locally in outputs/300/phylo

conda activate py2
~/github/mccloud-rrt/vcf2phylip.py -i pruned01-max.vcf
~/github/mccloud-rrt/vcf2phylip.py -i pruned01-rand.vcf

~/github/mccloud-rrt/vcf2phylip.py -i pruned05-max.vcf
~/github/mccloud-rrt/vcf2phylip.py -i pruned05-rand.vcf
~/github/mccloud-rrt/vcf2phylip.py -i pruned10-max.vcf
~/github/mccloud-rrt/vcf2phylip.py -i pruned10-rand.vcf

conda deactivate

iqtree2 -s pruned05-max.min4.phy -T AUTO -st DNA -m MFP+ASC -bb 10000 --redo
iqtree2 -s pruned05-max.min4.phy.varsites.phy -T AUTO -st DNA -m GTR+ASC -bb 1000 -nm 2000 -bcor 0.9 --redo

iqtree2 -s pruned05-rand.min4.phy -T AUTO -st DNA -m MFP+ASC -bb 10000 --redo
iqtree2 -s pruned05-rand.min4.phy.varsites.phy -T AUTO -st DNA -m GTR+ASC -bb 1000 -nm 2000 -bcor 0.9 --redo

iqtree2 -s pruned10-max.min4.phy -T AUTO -st DNA -m MFP+ASC -bb 10000 --redo
iqtree2 -s pruned10-max.min4.phy.varsites.phy -T AUTO -st DNA -m GTR+ASC -bb 1000 -nm 2000 -bcor 0.9 --redo

iqtree2 -s pruned10-rand.min4.phy -T AUTO -st DNA -m MFP+ASC -bb 10000 --redo
iqtree2 -s pruned10-rand.min4.phy.varsites.phy -T AUTO -st DNA -m GTR+ASC -bb 1000 -nm 2000 -bcor 0.9 --redo

```


```{r}

dat<-read.dna(file="outputs/300/pruned01-max.min4.phy")
write.nexus.data(dat, file="outputs/300/pruned01-max.nex")

dat<-read.dna(file="outputs/300/pruned01-rand.min4.phy")
write.nexus.data(dat, file="outputs/300/pruned01-rand.nex")

dat<-read.dna(file="outputs/300/phylo/pruned10-rand.min4.phy.varsites.phy")
write.nexus.data(dat, file="outputs/300/pruned-10-rand.nex")

dat<-read.dna(file="outputs/300/phylo/pruned10-max.min4.phy.varsites.phy")
write.nexus.data(dat, file="outputs/300/pruned-10-max.nex")

```

Checking for excessive missingness
```{sh, eval=FALSE}
paste \
<(bcftools query -f '[%SAMPLE\t]\n' erebi-filtered.vcf.gz | head -1 | tr '\t' '\n') \
<(bcftools query -f '[%GT\t]\n' erebi-filtered.vcf.gz | awk -v OFS="\t" '{for (i=1;i<=NF;i++) if ($i == "./.") sum[i]+=1 } END {for (i in sum) print i, sum[i] / NR }' | sort -k1,1n | cut -f 2) > missingness.stats
 cat missingness.stats | sort -k 2 -n | tail
```


# PCA
```{r}
path_to_file <- "outputs/300/pruned01-rand.vcf"

filename <- read.pcadapt(path_to_file, type = "vcf")

```


_1_ choose a K   

```{r}
x <- pcadapt(input = filename, K = 10) 
```

```{r}
var<-round(x$singular.values^2*100,2)
var
```

```{r}
#plot(x, option = "screeplot")
panela<-plot(x, option = "screeplot", K = 10) + ggtitle("A") +
  theme_bw() +
  theme(plot.title = element_text(size=14, face="bold")) +
  theme(panel.grid = element_blank(), panel.background = element_blank()) +
  scale_x_continuous(breaks=seq(1,10,1), labels=seq(1,10,1)) +
  ylab("Proportion of Explained Variance\n") +
  xlab("\nPrincipal Component") +
  theme(axis.title=element_text(size=12, face="bold"))
panela

ggsave("outputs/105/scree-plot.jpeg", width=4, height=4)
```


```{r}
plot(x, option = "scores")
```


```{r}
m<-meta
```

```{r}
pcadata<-x$scores[,1:5] %>% as_tibble()
pcadata<-pcadata %>% bind_cols(m)

```

```{r}
p1<-ggplot(pcadata) +
  geom_point(aes(x=V1, y=V2, fill=`Tributary/Collection Location`), pch=21, alpha=0.75, cex=2) +
  theme_bw() +
 # theme(legend.position = "top") +
  theme(panel.grid = element_blank()) +
  xlab("PC 1") +
  ylab("PC 2") +
  scale_fill_viridis_d(option="H") +
  xlab(paste0("PC1 ",var[1],"%")) +
  ylab(paste0("PC2 ",var[2],"%")) +
  theme(axis.title=element_text(size=12, face="bold"))


p1

```

# PCA
```{r}
path_to_file <- "outputs/300/newarkensis-pruned05-max.vcf"

filename <- read.pcadapt(path_to_file, type = "vcf")

```


_1_ choose a K   

```{r}
x <- pcadapt(input = filename, K = 10) 
```

```{r}
var<-round(x$singular.values^2*100,2)
var
```

```{r}
#plot(x, option = "screeplot")
panela<-plot(x, option = "screeplot", K = 10) + ggtitle("A") +
  theme_bw() +
  theme(plot.title = element_text(size=14, face="bold")) +
  theme(panel.grid = element_blank(), panel.background = element_blank()) +
  scale_x_continuous(breaks=seq(1,10,1), labels=seq(1,10,1)) +
  ylab("Proportion of Explained Variance\n") +
  xlab("\nPrincipal Component") +
  theme(axis.title=element_text(size=12, face="bold"))
panela

ggsave("outputs/105/scree-plot.jpeg", width=4, height=4)
```


```{r}
plot(x, option = "scores")
```


```{r}
m<-nmeta
```

```{r}
pcadata<-x$scores[,1:5] %>% as_tibble()
pcadata<-pcadata %>% bind_cols(m)
labels12<-pcadata %>% select(`Tributary/Collection Location`, V1, V2) %>% group_by(`Tributary/Collection Location`) %>% summarize(X=mean(V1), Y=mean(V2)) %>% filter(`Tributary/Collection Location` %in% c("NV19","NV18","NV17","NV12"))

labels13<-pcadata %>% select(`Tributary/Collection Location`, V1, V3) %>% group_by(`Tributary/Collection Location`) %>% summarize(X=mean(V1), Y=mean(V3)) %>% filter(`Tributary/Collection Location` %in% c("NV19","NV18","NV17","NV12","Cold Creek Reservoir"))
labels13
```

```{r}
p1<-ggplot(pcadata) +
  geom_point(aes(x=V1, y=V2, fill=`Tributary/Collection Location`), pch=21, alpha=0.75, cex=2) +
  geom_label_repel(data=labels12, aes(x=X, y=Y, label=`Tributary/Collection Location`), alpha=0.75) +
  theme_bw() +
 # theme(legend.position = "top") +
  theme(panel.grid = element_blank()) +
  xlab("PC 1") +
  ylab("PC 2") +
  scale_fill_viridis_d(option="H") +
  xlab(paste0("PC1 ",var[1],"%")) +
  ylab(paste0("PC2 ",var[2],"%")) +
  theme(axis.title=element_text(size=12, face="bold"))


p1 + facet_wrap(.~`Tributary/Collection Location`)

```


```{r}
p2<-ggplot(pcadata) +
  geom_point(aes(x=V1, y=V3, fill=`Tributary/Collection Location`), pch=21, alpha=0.75, cex=2) +
  geom_label_repel(data=labels13, aes(x=X, y=Y, label=`Tributary/Collection Location`), alpha=0.75) +
  theme_bw() +
  #theme(legend.position = "none") +
  theme(panel.grid = element_blank()) +
  scale_fill_viridis_d(option="H") +
  xlab(paste0("PC1 ",var[1],"%")) +
  ylab(paste0("PC3 ",var[3],"%")) + 
  guides(fill=guide_legend(override.aes=list(shape=21))) +
    theme(axis.title=element_text(size=12, face="bold"))


p2
```

```{r}
ggplot(pcadata) +
  geom_point(aes(x=V1, y=V4, fill=`Tributary/Collection Location`), pch=21, alpha=0.75, cex=2) +
  theme_bw() +
  #theme(legend.position = "none") +
  theme(panel.grid = element_blank()) +
  scale_fill_viridis_d(option="H") +
  xlab(paste0("PC1 ",var[1],"%")) +
  ylab(paste0("PC4 ",var[4],"%")) + 
  guides(fill=guide_legend(override.aes=list(shape=21))) +
    theme(axis.title=element_text(size=12, face="bold"))
```

```{r}
ggarrange(p1+theme(legend.position = "none"),p2, widths=c(1,1.6))

ggsave("outputs/300/newarkensis-pca.jpeg", width=10, height=4)
```

## Let's pull out outliers on PC1 and see what they are


```{r}
x <- pcadapt(filename, K = 4)
summary(x)
```

```{r}
plot(x , option = "manhattan")
```

```{r}
plot(x, option = "qqplot")
```

```{r}
hist(x$pvalues, xlab = "p-values", main = NULL, breaks = 50, col = "orange")
```

```{r}
plot(x, option = "stat.distribution")
```


```{R}
par(mfrow = c(2, 2))
for (i in 1:4)
  plot(x$loadings[, i], pch = 19, cex = .3, ylab = paste0("Loadings PC", i))
```    
```{r}
plot(x)
```

```{r}
padj <- p.adjust(x$pvalues,method="bonferroni")
alpha <- 0.05
outliers <- which(padj < alpha)
length(outliers)

snp_pc <- get.pc(x, outliers)
snp_pc %>% group_by(PC) %>% summarize(Count=n())

```

```{r}
padj <- p.adjust(x$pvalues,method="BH")
alpha <- 0.05
outliers <- which(padj < alpha)
length(outliers)

snp_pc <- get.pc(x, outliers)
snp_pc %>% group_by(PC) %>% summarize(Count=n())

```

get snps `gunzip -c outputs/105/pruned-05-max.vcf.gz| grep -v "^#" | cut -f1-3 | awk '{print $0"\t"NR}' > outputs/105/pruned-05-max-snp-numbers.tsv`


```{r}
snp_pc1<-snp_pc %>% filter(PC==1)
labels<-read_tsv("meta/lates-lgs-named.txt") %>% rename(Name=Chrom, Chrom=Chromosome)
labels
```


```{r}
snpnumbers<-read_tsv("outputs/105/pruned-05-max-snp-numbers.tsv", col_names = c("Chrom","Pos","Marker","SNP"))
snps<-left_join(as_tibble(snp_pc1), snpnumbers) 

snps %>% group_by(PC, Chrom) %>% summarize(Count=n()) %>% arrange(-Count)

snps %>% left_join(labels) 
```

How about a network?

