---
title: "302-alkali-lake"
output: html_document
date: "2026-05-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
mk<-read_csv("meta/full-meta.csv")  %>% mutate(Path2=ifelse(Dedup > 2e6, paste0("data/downsample/",Ind,".reduced.bam"),
                     Path)) %>% mutate(Taxon=Species) %>% filter(Species %in% c("obesus","snyderi","toikona"))

m2k<-mk  %>% filter(Dedup > 2e5)


m2k %>% select(Path) %>% write_csv("bamlists/m2k.bamlist", col_names = FALSE)
m2k %>% select(Path2) %>% write_csv("bamlists/m2k-downsample.bamlist", col_names = FALSE)

## downsample some of the fish

m4k<-read_csv("meta/obesus-200k-meta-edited.csv") %>% mutate(Path=paste0("data/align/",Ind,".sort.flt.bam")) %>% 
  mutate(Path2=ifelse(Dedup > 2e6, paste0("data/downsample/",Ind,".reduced.bam"),
                     Path)) %>% mutate(Taxon=Species) %>% filter(Dedup > 4e5) 
m4k$Species<-factor(m5k$Species, levels=c("newarkensis","isolatus","bicolor","thalassinus","mohavensis","snyderi","obesus"))

obs<-m4k %>% filter(Species=="obesus") %>% slice_sample(n=5, by=`Tributary/Collection Location`)
nobs<-m4k %>% filter(Species!="obesus") %>% slice_sample(n=10, by=Species)

sub4<-bind_rows(obs, nobs)
sub4 %>% group_by(Species, Vicinity) %>% summarize(Count=n())

sub4 %>% select(Path2) %>% write_csv("bamlists/sub4-downsample.bamlist", col_names = FALSE)

```

548 inds with doIbs for coverage
164 when downsampled 
testing initially with single chrom JAPDVR010000001.1

```{sh, eval=FALSE}
srun -p bmh -t 36:00:00 --mem=128G --nodes=1 --ntasks-per-node=1 --cpus-per-task=12  $HOME/angsd/angsd -P 12  \
-bam bamlists/m2k-downsample.bamlist -ref /home/maccamp/genomes/gila-orcutti/GCA_026230005.1_fGilOrc1.0.hap1_genomic.fna.gz \
-r JAPDVR010000001.1 \
-minInd 493 -minMapQ 10 -minQ 20 -GL 2 -doGLF 2 \
-doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doIBS 1 -doCounts 1 -doCov 1 -makeMatrix 1 -minMaf 0.01 \
-out outputs/302/m2k-ibs-90 >outputs/302/m2k-ibs-90.out 2> outputs/302/m2k-ibs-90.err &

srun -p bmh -t 12:00:00 --mem=128G --nodes=1 --ntasks-per-node=1 --cpus-per-task=12  $HOME/angsd/angsd -P 12  \
-bam bamlists/sub4-downsample.bamlist -ref /home/maccamp/genomes/gila-orcutti/GCA_026230005.1_fGilOrc1.0.hap1_genomic.fna.gz \
-r JAPDVR010000001.1 \
-minInd 148 -minMapQ 10 -minQ 20 -GL 2 -doGLF 2 \
-doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doIBS 1 -doCounts 1 -doCov 1 -makeMatrix 1 -minMaf 0.01 \
-out outputs/302/sub4-ibs-90 >outputs/302/sub4-ibs-90.out 2> outputs/302/sub4-ibs-90.err &


```


srun -p high -t 04:00:00 --mem=32G --nodes=1 $HOME/angsd/misc/NGSadmix -P 12 -likes outputs/104/mo5-ibs-90-high.beagle.gz -K 2 -minMaf 0.01 -o outputs/104/mo5-K2 &
srun -p high -t 05:00:00 --mem=32G --nodes=1 $HOME/angsd/misc/NGSadmix -P 12 -likes outputs/104/mo5-ibs-90-high.beagle.gz -K 3 -minMaf 0.01 -o outputs/104/mo5-K3 &
srun -p high -t 06:00:00 --mem=32G --nodes=1 $HOME/angsd/misc/NGSadmix -P 12 -likes outputs/104/mo5-ibs-90-high.beagle.gz -K 4 -minMaf 0.01 -o outputs/104/mo5-K4 &
srun -p high -t 06:00:00 --mem=32G --nodes=1 $HOME/angsd/misc/NGSadmix -P 12 -likes outputs/104/mo5-ibs-90-high.beagle.gz -K 5 -minMaf 0.01 -o outputs/104/mo5-K5 &
srun -p high -t 06:00:00 --mem=32G --nodes=1 $HOME/angsd/misc/NGSadmix -P 12 -likes outputs/104/mo5-ibs-90-high.beagle.gz -K 6 -minMaf 0.01 -o outputs/104/mo5-K6 &

```
