---
title: "network"
output: html_document
date: "2024-05-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

Copying network from Gadopsis paper and then modifying for Tui Chub

```{r}
library(tidyverse)
library(tanggle)
library(ape)
```

module load julia/1.10.2

Adding dependencies on farm

```{sh, eval=FALSE}
using Pkg
Pkg.add("PhyloNetworks")
import Pkg; Pkg.add("CSV")
import Pkg; Pkg.add("DataFrames")
import Pkg; Pkg.add("PhyloPlots")
```

## Making concordance factors

Olave and Meyer (2020) present a may to calculate concordance factors from a phylip formatted file for input into PhyloNetworks.  So, let's try that.

First I cloned https://github.com/melisaolave/SNPs2CF.
Required dependencies.    
install.packages("foreach")    
install.packages("doMC")     

```{r, warning=FALSE, message=FALSE}
source("~/github/SNPs2CF/functions_v1.6.R")
```

Need an imap file and need a phylip file. The imap file has Traits and Species. Traits is the sequence name and Species is the group.
Outgroups
G. orcuttii
S. alvordensis
P. macrolepidotus

Ingroups
isolatus
newarkensis (NN4, NV12 combined)
thalassinus vaccaceps (separate)
thalassinus  (combined locs)
bicolor columbianus Kueney County
bicolor eurysomas (Combined locs)
mohavensis Camp Cady, Tui Slough 
for toikona let's use Mule Spring / Cottonwood Pond combined
for snyderi let's use Sotcher Lake, NE Pond, SW Pond, SE Pond combined
for obesus 
Fish Lake Valley Lida Pond,
Fish Lake Valley McNett Ranch
Little Fish Lake Valley
Hot Creek Valley/Railroad Valley

Combined Lahontan lineage
Walker Lake/Rose Creek 
Eagle Lake 
Carson Desert, Little Soda Lake
Upper Humboldt

Need an imap file and need a phylip file. The imap file has Traits and Species. Traits is the sequence name and Species is the group.

Made one by hand with 21 species.  


```{r, warning=FALSE, message=FALE, eval=FALSE}
output <- SNPs2CF(seqMatrix="network/438-01.min4.phy",
                  ImapName="network/438.imap", between.sp.only = TRUE, #max.SNPs = 1000,
                  n.quartets = 3,
                  bootstrap=FALSE, outputName="network/network-nq3-btsp.csv",
                  save.progress=FALSE,
                  cores = 4);
```
Took 1068 seconds with 3 quartets, all snps and 21 species, 3 cores

## Credibility Intervals
To bootstrap I need to put together CF's with CIs

```{r, warning=FALSE, message=FALE, eval=FALSE}
ciout <- SNPs2CF(seqMatrix="network/438-01.min4.phy",
                  ImapName="network/438.imap", between.sp.only = TRUE, max.SNPs = NULL,
                  n.quartets = 3, bootstrap=TRUE, boots.rep=1000,
                  outputName="network/network-nq3-btsp-CI.csv",
                  save.progress=FALSE,
                  cores = 4);
```


# Starting tree
creating a net0 tree like so, veryfing it works

On farm

```{sh, eval=FALSE}
using PhyloNetworks;    
using CSV, DataFrames;     

CF=readTableCF("/home/maccamp/fish-lake/network/network-nq3-btsp.csv");    

using PhyloPlots;    
treefile = joinpath("/home/maccamp/fish-lake/network/tree.tre");
tree = readTopology(treefile);     
plot(tree, :R, showEdgeLength=true);

T=readTopologyLevel1(treefile);    

# Using snaq!
net0 = snaq!(T,CF, hmax=0, filename="/home/maccamp/fish-lake/outputs/107/net0-start", seed=1234);      
writeTopology(net0, "/home/maccamp/fish-lake/outputs/107/net0-start.tre")
```

next, running network.slurm and network.jl with hmax=5