---
title: "treemix"
output: html_document
date: "2024-05-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


```{r}
library(tidyverse)
```

Get pops for treemix

```{r}
treemix<-read_csv("meta/phylo-582.csv") %>% filter(!`Tributary/Collection Location` %in% c("Cowhead Lake","Desert Discovery Pond")) %>%
  filter(`Tributary/Collection Location` != "3/8 Mile Spring")
```

573 inds (less G. orcuttii P. macrolepidotus)
571 inds (less 3/8 spring)

Need a .clst file,
three fields no header
FamilyID WithinFamilyID ClusterName

Oh argghhhhh, need some clusters

```{r}
clst<-treemix %>% mutate(Inter=GVL_Code) %>% mutate(Cluster=paste0(Species,"_",`Tributary/Collection Location`)) 
clst$Cluster<-gsub(" ","_",clst$Cluster)

clst<-clst %>% filter(!Cluster %in% c("obesus_Little_Hot_Creek_Pond","obesus_Little_Hot_Creek","obesus_Upper_Owens_Gorge","obesus_SE_Pond_(White_Mountain_Research_Center",
                                      "obesus_SW_Pond_(White_Mountain_Research_Center")) %>% group_by(Cluster) %>% mutate(Count=n())
clst %>% filter(Count >2) %>% summarize(Count=n())

clst %>% dplyr::select(GVL_Code, Sample, Cluster) %>% write_tsv(file="treemix/treemix.clst", col_names = FALSE)

clst %>% write_csv("treemix/treemix-meta.csv")
write_csv(clst %>% ungroup() %>% dplyr::select(Sample), file="treemix/treemix.names", col_names=FALSE)

```

Can combine some of these (Eagle Lake, Flowing Wells), but I dumped small sizes.

550 individuals

```{sh, eval=FALSE}
bcftools  view -S treemix/treemix.names outputs/103/plink-01.vcf.gz | bcftools +fill-tags | bcftools view -q 0.01:minor | bcftools +prune -m 0.20 -w 10000 > treemix/treemix-pruned.vcf

bgzip treemix/treemix-pruned.vcf
tabix treemix/treemix-pruned.vcf.gz
```

git filter-branch --force --index-filter \
  'git rm --cached --ignore-unmatch treemix/treemix-pruned.vcf' \
  --prune-empty --tag-name-filter cat -- --all
  
Now I should be able to work this out on the other computer.